Association between SARS-CoV-2 and metagenomic content of samples from the Huanan Seafood Market

### 2023-07-02 paper/reviewer_response.tex

\response{
I agree that the calling of samples as positive / negative based on the binary presence / absence of any SARS-CoV-2 read is not statistically well supported, as we cannot reject the null hypothesis that the underlying amount of SARS-CoV-2 is significantly lower in a sample with 0 in 200-million SARS-CoV-2 reads versus a sample with 1 in 200-million SARS-CoV-2 reads.
If we use a Fisher exact test and compare two samples with 200-million total reads one of which has 0 SARS-CoV-2 reads and the other of which has $N$ SARS-CoV-2 reads, we can only reject the null hypothesis that the two samples are drawn from reads with the same underlying distribution of SARS-CoV-2 content at $P \le 0.05$ when $N \ge 6$.

Therefore, as suggested by the reviewer, I have expanded on the relevant portion of the text explaining this fact, and described what threshold is needed for statistical significance on the statement that the SARS-CoV-2 content of the two samples is significantly different.

That said, I have not included any claims about how this fact would affect the conclusions that have been drawn by Worobey and others.
Certainly it is a factor that needs to be considered---however, many of the samples analyzed by Worobey and co-workers were also analyzed by RT-qPCR in addition to next-generation sequencing, so substantially more analysis of the full dataset and spatial distribution of samples is needed to determine how the above point might affect any conclusions by Worobey et al.
That analysis is beyond the scope of the current study, so here I have simply noted the lack of statistical rigor in how samples have been called positive / negative from deep-sequencing data by the Chinese CDC, but have refrained from speculating on how exactly that might impact downstream analyses using the Chinese CDC's positive / negative classifications.
}

5. I think several figures can be removed from the printed manuscript without information loss. For example Figure 1 shows a lot of near ~1 correlations, and Figure 4 shows a lot of ``Rorschach blot'' scatter-shots of points.

\response{
As the reviewer suggests, I have moved Figure 1 of the original manuscript to the supplement, where it is now Figure S1.
This is the figure that simply shows that the mammalian metagenomic composition is highly correlated across all shared samples between my analysis and that of Crits-Christoph et al.

However, I have chosen to keep Figure 1 of the original manuscript (now Figure 3 of the revised manuscript) as a main figure.
I agree with the reviewer that this is a large figure with lots of scatter plots that is a bit overwhelming to look at.
But this is also represents arguably the most key results of the entire analysis, and since \textit{Virus Evolution} has no length limit I think it is better to keep this large full-page figure so all of the key results are in the main manuscript without having to refer to the supplement.
}

### 2023-07-03 paper/reviewer_response.tex

\response{
As the reviewer suggests, I have moved Figure 1 of the original manuscript to the supplement, where it is now Figure S1.
This is the figure that simply shows that the mammalian metagenomic composition is highly correlated across all shared samples between my analysis and that of Crits-Christoph et al.

However, I have chosen to keep Figure 1 of the original manuscript (now Figure 3 of the revised manuscript) as a main figure.
I agree with the reviewer that this is a large figure with lots of scatter plots that is a bit overwhelming to look at.
But this is also represents arguably the most key results of the entire analysis, and since \textit{Virus Evolution} has no length limit I think it is better to keep this large full-page figure so all of the key results are in the main manuscript without having to refer to the supplement.
}

6. For regression analyses, I might suggest something more robust to outliers that linear/rank regression. Theil-Sen estimators, or quantile regression might be better here.

\response{
The reviewer is correct that there are a lot of ``outlier'' points in the correlation scatter plots.
This isn't a substantive issue for the large panel of scatter plots in Figure 3 of the revised manuscript (Figure 4 of the original) since the panels show all the points visually and don't rely on the correlation coefficients for anything other than a numerical summary.
However, this issue does substantively affect Figure 4 of the revised manuscript (Figure 3 of the original), which plots the Pearson and Spearman correlation coefficients.

The reviewer is correct that Theil-Sen estimators or quantile regression are more robust to outliers.
However, there are two difficulties with using these instead: the paper never shows the fit lines (which is what those approaches estimate), but rather the correlation coefficient.
But both Theil-Sen estimators nor quantile regression are designed to estimate fit lines, and don't directly estimate a counterpart to a correlation coefficient.
However, I did locate a formula for a correlation measure based on the Theil-Sen estimator (\url{https://towardsdatascience.com/a-correlation-measure-based-on-theil-sen-regression-31b8b9ed64f1}).
I have therefore created a version of Figure 4 that uses this Theil-Sen estimator based correlation rather than the Pearson correlation; it is Figure S4 of the revised manuscript.
The correlations are somewhat different for the Theil-Sen based estimator than the Pearson or Spearman methods (the Theil-Sen measure is zero for many species where there is no correlation except for a few outliers), but the basic finding holds that the most correlated species across the entire dataset do not include non-human species susceptible to SARS-CoV-2.

I have kept the Pearson / Spearman correlation figure as the main text figure for consistency with the Chinese CDC pre-print, and put this new plot as a supplementary figure.
}

### 2023-07-03 paper/paper.tex
Reviewer 2 comment 3: move figure to supplement
Reviewer 2 comment 4: Theil-Sen correlation

\begin{figure}
\includegraphics[width=\linewidth]{figures/overall_corr/overall_corr.pdf}
\caption{
Correlations between SARS-CoV-2 content and mitochondrial content for all species.
Top row show just samples containing at least one SARS-CoV-2 read and bottom row shows samples regardless of SARS-CoV-2 content; left shows samples from all sampling dates and right shows just samples from the January-12-2020 date when most of the wildlife sampling occurred.
This plot is designed to mimic the fourth figure of \citet{liu2022surveillance}.
Some key species are labeled; see the interactive version at \url{https://jbloom.github.io/Huanan_market_samples/overall_corr.html} to mouseover all points for details, select different subsets of samples, and calculate the correlations on a linear or log scale.
\label{fig:overall_corr}
}
\end{figure}

To remedy this omission, I used the chordate mitochondrial compositions calculated in the current study to generate plots analogous to that in the original Chinese CDC preprint.
When the correlations are taken across all sampling dates only for samples containing at least one SARS-CoV-2 read (as done in the original Chinese CDC preprint), then the genetic material of several non-human species (largemouth bass, catfish, cow, sheep, pig) is more correlated with SARS-CoV-2 reads than is human material (Figure~\ref{fig:overall_corr}, upper left).
If the plot is expanded to all samples (regardless of whether or not they contain SARS-CoV-2 reads), the trends are broadly similar, with fish remaining the most correlated species (Figure~\ref{fig:overall_corr}, lower left).
It is also informative to take the correlations only across samples collected on January-12-2020, which was the date of most intensive sampling of the wildlife stalls (this is when Q61 was collected).
 For January-12-2020 samples that contain SARS-CoV-2 reads, humans are among the species whose genetic material is most correlated with SARS-CoV-2 reads, but goat and spotted dove have roughly equivalent correlations (Figure~\ref{fig:overall_corr}, upper right).
 If we include January-12-2020 samples regardless of SARS-CoV-2 content, then snakehead fish and Malayan porcupine are among the most correlated species.
Raccoon dog and bamboo rat genetic material are not positively correlated with SARS-CoV-2 reads in any of the sample sets.
See the interactive version of Figure~\ref{fig:overall_corr} at \url{https://jbloom.github.io/Huanan_market_samples/overall_corr.html} to explore other sample subsets.

@@

>>>>

\begin{figure}
\includegraphics[width=\linewidth]{figures/overall_corr/overall_corr.pdf}
\caption{
Correlations between SARS-CoV-2 content and mitochondrial content for all species.
Top row show just samples containing at least one SARS-CoV-2 read and bottom row shows samples regardless of SARS-CoV-2 content; left shows samples from all sampling dates and right shows just samples from the January-12-2020 date when most of the wildlife sampling occurred.
This plot is designed to mimic the fourth figure of \citet{liu2022surveillance}.
Some key species are labeled; see the interactive version at \url{https://jbloom.github.io/Huanan_market_samples/overall_corr.html} to mouseover all points for details, select different subsets of samples, and calculate the correlations on a linear or log scale.
See Figure~\ref{fig:theil_sen_corr} for a version of this plot that uses correlations based on a Theil-Sen estimator that is more robust to outliers.
\label{fig:overall_corr}
}
\end{figure}

To remedy this omission, I used the chordate mitochondrial compositions calculated in the current study to generate plots analogous to that in the original Chinese CDC preprint.
When the correlations are taken across all sampling dates only for samples containing at least one SARS-CoV-2 read (as done in the original Chinese CDC preprint), then the genetic material of several non-human species (largemouth bass, catfish, cow, sheep, pig) is more correlated with SARS-CoV-2 reads than is human material (Figure~\ref{fig:overall_corr}, upper left).
If the plot is expanded to all samples (regardless of whether or not they contain SARS-CoV-2 reads), the trends are broadly similar, with fish remaining the most correlated species (Figure~\ref{fig:overall_corr}, lower left).
It is also informative to take the correlations only across samples collected on January-12-2020, which was the date of most intensive sampling of the wildlife stalls (this is when Q61 was collected).
 For January-12-2020 samples that contain SARS-CoV-2 reads, humans are among the species whose genetic material is most correlated with SARS-CoV-2 reads, but goat and spotted dove have roughly equivalent correlations (Figure~\ref{fig:overall_corr}, upper right).
 If we include January-12-2020 samples regardless of SARS-CoV-2 content, then snakehead fish and Malayan porcupine are among the most correlated species.
Raccoon dog and bamboo rat genetic material are not positively correlated with SARS-CoV-2 reads in any of the sample sets.
See the interactive version of Figure~\ref{fig:overall_corr} at \url{https://jbloom.github.io/Huanan_market_samples/overall_corr.html} to explore other sample subsets.
The trends are broadly similar if the correlations are instead calculated based on a Theil-Sen estimator~\citep{theilsen} that is more robust to outliers (Figure~\ref{fig:theil_sen_corr}).

@@

\begin{figure}
\centering
\includegraphics[width=0.6\linewidth]{figures/theil_sen_corr/theil_sen_corr.png}
\caption{
A version of Figure~\ref{fig:overall_corr} where the y-axis shows a correlation computed from the Theil-Sen estimator~\citep{theilsen} rather than the Pearson correlation.
The Theil-Sen estimator is more robust to outliers, so species where the correlation is driven by just a few species will have values of zero in these plots.
See \url{https://jbloom.github.io/Huanan_market_samples/theil_sen_corr.html} for an interactive version of this plot.
\label{fig:theil_sen_corr}
}
\end{figure}
